---
title: "DrugResponsePredictions"
author: 
- "Vigneshwari Subramanian"
-  "Attila Gabor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```

Generate drug response prediction models using cell line (Mutation / CNV) and Drug features (ECFP4 descriptors)
The function used here is meant for LNIC50 predictions. If some other metric (GR50, AUC, GRAOC) is used, it should be changed accordingly in the function.

For practical reasons, Model building and predictions are not done locally and therefore, the results are not shown in R markdown.


```{r warnings=FALSE, suppressMessages=TRUE}
#Load packages
if(!(require(c("caret","Metrics","AUC","car","proxy")))) {install.packages(c("caret","Metrics","AUC","car","proxy","doMC","randomForest"))}
p<-c("caret","Metrics","AUC","car","proxy","doMC","randomForest")
lapply(p,require,character.only=TRUE)
```

Load data and generate drug response prediction models

```{r}
data("LigandDesc_LNIC50")
data("MutCNV")
DrugResponse_pred <- DrugResponsePredictions_Ml_models(LigandDesc_LNIC50,MutCNV )
```

