---
title: "Alternative Metrics for GDSC"
author: 
- "Vigneshwari Subramanian"
-  "Attila Gabor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Alternative Metrics for GDSC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Rscript with functions to calculate GR and all associated metrics based on the approach described by Marc Hafner in \url{https://www.nature.com/articles/nbt.3882}


```{r load_AlternativeDrugResponseMetrics}
library(AlternativeDrugResponseMetrics)
```
Install and load a few packages required for this analysis

```{r message = FALSE, warning = FALSE}
if(!(require(c("drc","reshape2","VennDiagram"),character.only = TRUE))) {install.packages(c("drc","reshape2","VennDiagram"))}
library("drc")
library("reshape2")
library("VennDiagram")
```

Data required: Raw data file with information about concentrations,raw intensities, positive and negative controls

Data downloaded from \url{ftp://ftp.sanger.ac.uk/pub/project/cancerrxgene/releases/current_release/v17.3_public_raw_data.csv} and Drug names mapped based on Drug IDs 

Read the raw data with the mapped drug names from the package. 
you can check the format of the input file by ```??AllDrugs_CellLines_GDSC```

```{r}
data("AllDrugs_CellLines_GDSC",package = "AlternativeDrugResponseMetrics")
GDSC_Data = as.data.frame(AllDrugs_CellLines_GDSC)
```


Compute the concetrations for each cellline - drug experiment:
```{r, include=FALSE}
GDSC_allvia = preprocess_GDSC(RawData = GDSC_Data,writeFile = T )
head(GDSC_allvia)
```


Load the Division times and viabilities from Genentech screening data (GCSI) 
```{r}
data("Genentech_GCSI_Gemcitabine_Response",package = "AlternativeDrugResponseMetrics")
Gemcitabine_GCSI_complete<-Genentech_GCSI_Gemcitabine_Response[complete.cases(Genentech_GCSI_Gemcitabine_Response),]

data("Genentech_GCSI_DivisionTimes_CellLines")
```


Extract the Gemcitabine data corresponding to those cell lines for which the division rates are known.
Make sure that the cell lines are in the same order in viabilities and divsion rates dataframes and  
```{r}
Cells_DivisionTimes_extract <- Genentech_GCSI_DivisionTimes_CellLines[ Genentech_GCSI_DivisionTimes_CellLines$CellLine %in% Gemcitabine_GCSI_complete$CellLine, ]
Cells_DivisionTimes_extract_ordered <- Cells_DivisionTimes_extract[ order(Cells_DivisionTimes_extract$CellLine), ]
Gemcitabine_DivisionTimes_known <- Gemcitabine_GCSI_complete[ Gemcitabine_GCSI_complete$CellLine %in% Genentech_GCSI_DivisionTimes_CellLines$CellLine, ]
Gemcitabine_GCSI_complete_ordered<-Gemcitabine_DivisionTimes_known[order(Gemcitabine_DivisionTimes_known$CellLine),]
Assay_Time <- 72
```


Calculate GR for Gemcitabine by using GCSI data
Extract Gemcitabine measurements made at concentrations between 0.1 and 32 Mm for calculating interquartile means; Done to avoid using noisy data for calculating GRref/ICrefs from Gemcitabine   
```{r}
Gemcitabine_GCSI_GR <- GCSI_GR( Gemcitabine_GCSI_Via_Doses = Gemcitabine_GCSI_complete_ordered,
								CellDivisionRates = Cells_DivisionTimes_extract_ordered,
								Assay_Time = Assay_Time,
								writeFile = T)
Gemcitabine_GR_filtered <- Gemcitabine_GCSI_GR[ which( Gemcitabine_GCSI_GR$Conc >= 0.1 & Gemcitabine_GCSI_GR$Conc <= 32), ]
Gemcitabine_GR_filtered_Lt32 <- Gemcitabine_GCSI_GR[ which(Gemcitabine_GCSI_GR$Conc <= 32), ]
```

Calculate Gemcitabine GRRef
```{r}
Gemcitabine_GCSI_InterquartileMean <- InterquartileMean(Gemcitabine_GR_filtered, type = "GCSI_GRRef", metric = "GR")
```

Calculate Gemcitabine ICRef
```{r}
GDSC_allvia_complete<-GDSC_allvia[complete.cases(GDSC_allvia),]

#Drug id corresponding to Gemcitabione in GDSC : 135
GDSC_Gemcitabine_extract<-GDSC_allvia_complete[GDSC_allvia_complete$Drug=="135",]
Gemcitabine_conc_filt<-GDSC_Gemcitabine_extract[which(GDSC_Gemcitabine_extract$Conc>=0.1 & GDSC_Gemcitabine_extract$Conc<=32),]
Gemcitabine_conc_filt_Lt32<-GDSC_Gemcitabine_extract[which(GDSC_Gemcitabine_extract$Conc<=32),]
Gemcitabine_GDSC_InterquartileMean<-InterquartileMean(Gemcitabine_GR_filtered = Gemcitabine_conc_filt, type = "GDSC_ICRef", metric = "Rel_via")
```

Load the already published GR and conventional metrics for GCSI data and extract Gemcitabine data
```{r}
data("Genentech_GCSI_GR_published",package = "AlternativeDrugResponseMetrics")
GCSI_Gemcitabine_published<-Genentech_GCSI_GR_published[Genentech_GCSI_GR_published$Perturbagen=="gemcitabine",]
```


Filter data to reduce noise and pick the cell lines for further analysis
```{r warning=FALSE}
Gemcitabine_R2_IC50_computed<-SigmoidFit_ComputeR2_IC50(Gemcitabine_conc_filt_Lt32)
Gemcitabine_GR_var<-CellLines_var(Gemcitabine_GR_filtered,"GR")
Gemcitabine_via_var<-CellLines_var(Gemcitabine_conc_filt,"Rel_via")
GDSC_finalCellLines<-CellLines_filter(Gemcitabine_GCSI_InterquartileMean,Gemcitabine_GR_var,Gemcitabine_GDSC_InterquartileMean,Gemcitabine_via_var,Gemcitabine_R2_IC50_computed,GCSI_Gemcitabine_published)
```




Compute GR for GDSC
```{r}
GDSC_GR_all<-GDSC_GR(GDSC_finalCellLines = GDSC_finalCellLines,IGDSC_ref = Gemcitabine_GDSC_InterquartileMean, GR_GCSI_ref = Gemcitabine_GCSI_InterquartileMean, GDSC_via_complete = GDSC_allvia_complete,writeFile = T )
```



Fit sigmoid curves to compute all the associated metrics(It takes ~30 Minutes for computation)
```{r warning=FALSE}
Metrics_GR_headers=cbind.data.frame("CellLine","Drug","GR50","GEC50","GRmax","GR_inf","GR_HillCoeff","GRAOC","GR_R2","LN_GR50")
GR_allMetrics<-AllMetrics_Calc(GDSC_GR_all,"GR",Metrics_GR_headers,writeFile=T,"Calculated_values/AllGDSC_ComputedGRMetrics_ConcCapped.txt")
colnames(GR_allMetrics)<-c("CellLine","Drug","GR50","GEC50","GRmax","GR_inf","GR_HillCoeff","GRAOC","GR_R2","LN_GR50")
IC_allMetrics<-ICMetrics_Calc(GDSC_GR_all,writeFile=T,"Calculated_values/AllGDSC_ComputedICMetrics_ConcCapped.txt")
colnames(IC_allMetrics)<-c("CellLine","Drug","IC50","EC50","Emax","IC_inf","IC_HillCoeff","AUC","IC_R2","LN_IC50","AAC")
```


Combine conventional and GR metrics; Pick only those cell line and drug pairs that have both IC and GR metrics
```{r}
Conventional_GR_allCalculatedMetrics_merged<-merge(GR_allMetrics,IC_allMetrics,by = c("CellLine","Drug"))
save(Conventional_GR_allCalculatedMetrics_merged, file = "Conventional_GR_allCalculatedMetrics_merged.rda")
```

Prepare the input files suitable for ANOVA analysis using GDSC tools (Conventional and GR metrics)
```{r}
data("COSMIC_ID_all",package = "AlternativeDrugResponseMetrics")
data("GDSC_orig_IC50_AUC",package = "AlternativeDrugResponseMetrics")
data("Conventional_GR_allCalculatedMetrics_merged",package="AlternativeDrugResponseMetrics")
IC50_GDSCTools_format<-GDSCTools_Features_prep(Conventional_GR_allCalculatedMetrics_merged,COSMIC_ID_all,"LN_IC50","_GRset")
AUC_GDSCTools_format<-GDSCTools_Features_prep(Conventional_GR_allCalculatedMetrics_merged,COSMIC_ID_all,"AUC","_GRset")
AAC_GDSCTools_format<-GDSCTools_Features_prep(Conventional_GR_allCalculatedMetrics_merged,COSMIC_ID_all,"AAC","_GRset")
GR50_GDSCTools_format<-GDSCTools_Features_prep(Conventional_GR_allCalculatedMetrics_merged,COSMIC_ID_all,"LN_GR50","_GRset")
GRAOC_GDSCTools_format<-GDSCTools_Features_prep(Conventional_GR_allCalculatedMetrics_merged,COSMIC_ID_all,"GRAOC","_GRset")
GRmax_GDSCTools_format<-GDSCTools_Features_prep(Conventional_GR_allCalculatedMetrics_merged,COSMIC_ID_all,"GRmax","_GRset")
Emax_GDSCTools_format<-GDSCTools_Features_prep(Conventional_GR_allCalculatedMetrics_merged,COSMIC_ID_all,"Emax","_GRset")
Orig_IC50_GDSCTools_format<-GDSCTools_Features_prep(GDSC_orig_IC50_AUC,COSMIC_ID_all,"LN_IC50","_GRset")
Orig_AUC_GDSCTools_format<-GDSCTools_Features_prep(GDSC_orig_IC50_AUC,COSMIC_ID_all,"AUC","_GRset")
```


Calculate NDR, PI and GR metrics based on the approach described in Gupta et al., 2018; The dataset being large and all the metrics being computed at the same time, it takes quite a few hours for computation
```{r}
#Concentrations and relative viabilities calculated during the initial preprocessing step; Filter the raw data to pick those cell lines with division times
data("AllDrugs_CellLines_GDSC",package = "AlternativeDrugResponseMetrics")
data("Genentech_GCSI_DivisionTimes_CellLines",package = "AlternativeDrugResponseMetrics")
GDSC_selective_data<-AllDrugs_CellLines_GDSC[AllDrugs_CellLines_GDSC$CellLine %in% Genentech_GCSI_DivisionTimes_CellLines$CellLine,]

#Compute NDR, PI and other metrics
RelVia_GR_NDR_PImetrics<-NDR_PI_GRmetrics(GDSC_selective_data,Genentech_GCSI_DivisionTimes_CellLines,writeFile=T)
colnames(RelVia_GR_NDR_PImetrics)<-c("CellLine","Drug","Conc","Via","NDR","PI","GR")
head(RelVia_GR_NDR_PImetrics)
```

Calculate all associated metrics for GR, PI and NDR data
```{r, warning=FALSE}
data(RelVia_GR_NDR_PImetrics)
Metrics_GR_headers=cbind.data.frame("CellLine","Drug","GR50","GEC50","GRmax","GR_inf","GR_HillCoeff","GRAOC","GR_R2","LN_GR50")
Metrics_NDR_headers=cbind.data.frame("CellLine","Drug","NDR50","NDR_EC50","NDRmax","NDR_inf","NDR_HillCoeff","NDR_AOC","NDR_R2","LN_NDR_50")
Metrics_PI_headers=cbind.data.frame("CellLine","Drug","PI50","PI_EC50","PImax","PI_inf","PI_HillCoeff","PI_AOC","PI_R2","LN_PI_50")

#Fit sigmoid curves to compute all related metrics
GR_allMetrics_349CellLines<-AllMetrics_Calc(AllOtherMetrics,"GR",Metrics_GR_headers,writeFile=T,"Calculated_values/AllGDSC_ComputedGRMetrics_349CellLines_ConcCapped.txt")
colnames(GR_allMetrics_349CellLines)<-c("CellLine","Drug","GR50","GEC50","GRmax","GR_inf","GR_HillCoeff","GRAOC","GR_R2","LN_GR50")
NDR_allMetrics_349CellLines<-AllMetrics_Calc(AllOtherMetrics,"NDR",Metrics_NDR_headers,writeFile=T,"Calculated_values/AllGDSC_ComputedNDRMetrics_349CellLines_ConcCapped.txt")
colnames(NDR_allMetrics_349CellLines)<-c("CellLine","Drug","NDR50","NDR_EC50","NDRmax","NDR_inf","NDR_HillCoeff","NDR_AOC","NDR_R2","LN_NDR_50")
PI_allMetrics_349CellLines<-AllMetrics_Calc(AllOtherMetrics,"PI",Metrics_PI_headers,writeFile=T,"Calculated_values/AllGDSC_ComputedPIMetrics_349CellLines_ConcCapped.txt")
colnames(PI_allMetrics_349CellLines)<-c("CellLine","Drug","PI50","PI_EC50","PImax","PI_inf","PI_HillCoeff","PI_AOC","PI_R2","LN_PI_50")
IC_allMetrics_349CellLines<-ICMetrics_Calc(AllOtherMetrics,writeFile=T,"Calculated_values/AllGDSC_ComputedICMetrics_349CellLines_ConcCapped.txt")
colnames(IC_allMetrics_349CellLines)<-c("CellLine","Drug","IC50","EC50","Emax","IC_inf","IC_HillCoeff","AUC","IC_R2","LN_IC50","AAC")
```


Combine conventional and other metrics; Pick only those cell line and drug pairs that have values for IC, GR, NDR and PI metrics
```{r}
library(plyr)
l_NDRSet_files<-list(GR_allMetrics_349CellLines,NDR_allMetrics_349CellLines,PI_allMetrics_349CellLines,IC_allMetrics_349CellLines)
AllMetrics_merged_NDRset<-join_all(l_NDRSet_files, type='inner', by = c("CellLine","Drug"))
save(AllMetrics_merged_NDRset,file="AllMetrics_merged_NDRset.rda")
```

Prepare the input files suitable for ANOVA analysis using GDSC tools (Conventional,GR,NDR and PI metrics)
```{r}
data("COSMIC_ID_all",package = "AlternativeDrugResponseMetrics")
data("AllMetrics_merged_NDRdata",package = "AlternativeDrugResponseMetrics")
IC50_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"LN_IC50","_NDRset")
AUC_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"AUC","_NDRset")
AAC_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"AAC","_NDRset")
GR50_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"LN_GR50","_NDRset")
GRAOC_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"GRAOC","_NDRset")
NDR50_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"LN_NDR_50","_NDRset")
NDRAOC_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"NDR_AOC","_NDRset")
PI50_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"LN_PI_50","_NDRset")
PIAOC_GDSCTools_format<-GDSCTools_Features_prep(AllMetrics_merged_NDRset,COSMIC_ID_all,"PI_AOC","_NDRset")
```
